---
title: "COVID19_round3"
author: "MELISSA"
date: "28/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries:
```{r}
library(Seurat)
library(base)
```

STEP1: Load the datasets in an automatic way:
```{r}
setwd("C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Controls_FilteredMatrices/")

tmp1=list.files("C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Controls_FilteredMatrices/")
#easynames=gsub(".txt.gz","",tmp1)

for(i in seq(1, length(tmp1))){
  assign(paste0("mat_", tmp1[i]),
         CreateSeuratObject( Read10X(paste0("C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Controls_FilteredMatrices/",tmp1[i], "/")), project = tmp1[i],min.cells = 20, min.features = 50))}
```

```{r}
nm=grep("mat_",ls(),value = T)
 for(i in seq(1, length(nm))){
   x=get(nm[i])
x[["percent.mito"]] <- PercentageFeatureSet(x, pattern = "^MT-")}
rm(x)
rm(easynames)
rm(i)
rm(nm)
rm(tmp1)
```

```{r}
mat_CTRL_02022[["percent.mito"]] <- PercentageFeatureSet(mat_CTRL_02022, pattern = "^MT-")
mat_CTRL_02021[["percent.mito"]] <- PercentageFeatureSet(mat_CTRL_02021, pattern = "^MT-")
mat_CTRL_02020RPMI[["percent.mito"]] <- PercentageFeatureSet(mat_CTRL_02020RPMI, pattern = "^MT-")
```

```{r}
library(ggplot2)
path="C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/"

p=VlnPlot(mat_CTRL_02020RPMI, c("nCount_RNA", "nFeature_RNA", "percent.mito"), combine = F)
p=VlnPlot(mat_CTRL_02021, c("nCount_RNA", "nFeature_RNA", "percent.mito"), combine = F)
p=VlnPlot(mat_CTRL_02022, c("nCount_RNA", "nFeature_RNA", "percent.mito"), combine = F)
####
for(i in 1:length(p)) {
  p[[i]] <- p[[i]] +NoLegend()+
   theme(axis.text.x = element_text(angle = 360))}
g1=cowplot::plot_grid(plotlist = p, ncol = 3)  
##
ggsave(filename=paste0("VlnPlot_CTRLEDTA2_",Sys.Date(),".png"), plot=g1, device="png",
       path= path, height=5, width=10, units="in", dpi=500) 

```

```{r}
seu= CreateSeuratObject(Read10X("/media/msaichi/MyPassport/Sauvegarde09112020/scRNAseq_COVID-19_round3/CellRanger_FilteredMatrices/01-051/filtered_feature_bc_matrix/"),min.cells = 20, min.features = 50)
seu[["percent.mito"]] <- PercentageFeatureSet(seu, pattern = "^MT-")
seu=subset(seu, subset  = nFeature_RNA < 6000 & nCount_RNA <40000 & percent.mito <25)
seu= preprocess(seu)

for (i in seq(1:length(ct))) {
 seu= AddModuleScore(seu, features = ct[i], name = cn[i])}

for (i in seq(1:length(ct))) {
 seu= AddModuleScore(seu, features = ct[i], name = cn[i])}
```

```{r}
 dim(mat_CTRL_02020RPMI)  15683  9884
```
```{r}
 dim(mat_CTRL_02021)  18643 38152
```
```{r} 
 dim(mat_CTRL_02022)  18312 27052
```
```{r}
mat_CTRL_02020RPMI=subset(mat_CTRL_02020RPMI, subset  = nFeature_RNA < 6000 & nCount_RNA <40000 & percent.mito <25)
mat_CTRL_02021=subset(mat_CTRL_02021, subset  = nFeature_RNA < 6000 & nCount_RNA <40000 & percent.mito <25)

mat_CTRL_02022=subset(mat_CTRL_02022, subset  = nFeature_RNA < 6000 & nCount_RNA <40000 & percent.mito <25)
```


```{r}
preprocess= function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 3000)
    x= ScaleData(x)
    x= RunPCA(x, features = VariableFeatures(object = x))
     x <- FindNeighbors(x, dims = 1:30)
    x <- FindClusters(x, resolution = 0.8)
   x<- RunUMAP(x, dims = 1:30)
}

mat_CTRL_02020RPMI= preprocess(mat_CTRL_02020RPMI)
mat_CTRL_02021= preprocess(mat_CTRL_02021)
mat_CTRL_02022= preprocess(mat_CTRL_02022)
```

```{r}
NK=list(c("NCAM1", "FCGR3A", "GNLY", "XCL1", "XCL2", "NCR1", "NKG7")) 
B=list(c("CD19", "MS4A1", "CD79A", "CD79B"))
plasmaCells=list(c("IGHG2", "IGHG1", "IGLC2", "IGHA1", "IGHA2", "IGHA3", "JCHAIN", "IGHM", "XBP1", "MZB1", "CD38", "IGLL5")) 
Tcells=list(c("CD3D", "CD3E", "CD3G"))
Platelet=list(c("PPBP"))
#######•
pDC=list(c("TCF4", "CLEC4C", "IRF7", "IRF8", "LILRA4", "IL3RA", "TLR9", "SPIB"))
cDC=list(c("ANPEP",  "CD1C", "ITGAX", "CST3", "FCER1A"))
monocytes=list(c("CD14", "FCGR1A"," S100A12", "FCGR3A", "MS4A7", "LYZ", "CXCR3"))
asDC=list(c("AXL", "SIGLEC6","CD22"))
  ct= c( pDC, cDC, monocytes, asDC)
cn= c( "pDC", "cDC", "monocytes", "asDC")

for (i in seq(1:length(ct))) {
 mat_CTRL_02020RPMI= AddModuleScore(mat_CTRL_02020RPMI, features = ct[i], name = cn[i])
 mat_CTRL_02021= AddModuleScore(mat_CTRL_02021, features = ct[i], name = cn[i])
mat_CTRL_02022= AddModuleScore(mat_CTRL_02022, features = ct[i], name = cn[i])}
######
c1=mat_CTRL_02020RPMI
c2=mat_CTRL_02021
c3=mat_CTRL_02022
c4=mat_CTRL_02020RPMI
```

```{r}
ct= c( NK, B, plasmaCells,Tcells, Platelet )
cn= c(  "NK", "B", "plasmaCells","Tcells", "Platelet")
for (i in seq(1:length(ct))) {
 c1= AddModuleScore(c1, features = ct[i], name = cn[i])
  c2= AddModuleScore(c2, features = ct[i], name = cn[i])
  c3= AddModuleScore(c3, features = ct[i], name = cn[i])
   c4= AddModuleScore(c4, features = ct[i], name = cn[i])}

keepAPC=function(x){
  x=subset(x, pDC1>0.2 | monocytes1>0.2 | cDC1>0.2 | asDC1>0.2) %>% subset(., NK1 <0 & B1 <0 & plasmaCells1 < 0 & Tcells1 < 0 )
  return(x)}
c1= keepAPC(c1) %>%  DietSeurat() %>% preprocess()
c2=keepAPC(c2) %>%DietSeurat(c2) %>% preprocess()
c3=keepAPC(c3) %>%DietSeurat(c3) %>% preprocess()

```
save the controls:
```{r}
saveRDS(c1, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/CTRL_02020RPMI_APC.Rds")
saveRDS(c2, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/CTRL_02021_APC.Rds")
saveRDS(c3, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/CTRL_02022_APC.Rds")
##
c1=readRDS("C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/CTRL_APC_ROUND3/CTRL_02020RPMI_APC.Rds")
c2= readRDS("C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/CTRL_APC_ROUND3/CTRL_02021_APC.Rds")
c3= readRDS("C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/CTRL_APC_ROUND3/CTRL_02022_APC.Rds")
```


#####################
LOAD THE PATIENTS:

```{r}
tmp1=list.files("C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Patients_FilteredMatrices/")
#easynames=gsub(".txt.gz","",tmp1)

for(i in seq(1, length(tmp1))){
  assign(paste0("P_", tmp1[i]),
         CreateSeuratObject( Read10X(paste0("C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Patients_FilteredMatrices/",tmp1[i], "/")), project = tmp1[i],min.cells = 20, min.features = 50))}
```

Add the mito genes ratio:
```{r}
nm=grep("P_",ls(),value = T)
P_01_047[["percent.mito"]] <- PercentageFeatureSet(P_01_047, pattern = "^MT-")
P_01_048_J0[["percent.mito"]] <- PercentageFeatureSet(P_01_048_J0, pattern = "^MT-")
P_01_048_J5[["percent.mito"]] <- PercentageFeatureSet(P_01_048_J5, pattern = "^MT-")
P_01_049[["percent.mito"]] <- PercentageFeatureSet(P_01_049, pattern = "^MT-")

P_01_050[["percent.mito"]] <- PercentageFeatureSet(P_01_050, pattern = "^MT-")
P_01_051[["percent.mito"]] <- PercentageFeatureSet(P_01_051, pattern = "^MT-")
P_01_051_J10[["percent.mito"]] <- PercentageFeatureSet(P_01_051_J10, pattern = "^MT-")
P_01_052[["percent.mito"]] <- PercentageFeatureSet(P_01_052, pattern = "^MT-")
P_01_053[["percent.mito"]] <- PercentageFeatureSet(P_01_053, pattern = "^MT-")
P_01_054_J0[["percent.mito"]] <- PercentageFeatureSet(P_01_054_J0, pattern = "^MT-")
```

```{r}
library(ggplot2)
path="C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/"

p=VlnPlot(P_01_047, c("nCount_RNA", "nFeature_RNA", "percent.mito"), combine = F)

QC=function(x){
  x=subset(x, subset  = nFeature_RNA < 6000 & nCount_RNA <40000 & percent.mito <25)
  return(x)}
P_01_047=QC(P_01_047)
P_01_048_J0=QC(P_01_048_J0)
P_01_048_J5=QC(P_01_048_J5)
P_01_049=QC(P_01_049)
P_01_050=QC(P_01_050)
P_01_051=QC(P_01_051)
P_01_051_J10= QC(P_01_051_J10)
P_01_052= QC(P_01_052)
P_01_053=QC(P_01_053)
P_01_054_J0= QC(P_01_054_J0)
####
  ctp= c( pDC, cDC, monocytes, asDC)
cnp= c( "pDC", "cDC", "monocytes", "asDC")

for (i in seq(1:length(ctp))) {
P_01_047= AddModuleScore(P_01_047, features = ctp[i], name = cnp[i])
P_01_048_J0= AddModuleScore(P_01_048_J0, features = ctp[i], name = cnp[i])
P_01_048_J5= AddModuleScore(P_01_048_J5, features = ctp[i], name = cnp[i])
P_01_049 = AddModuleScore(P_01_049, features = ctp[i], name = cnp[i])
P_01_050 = AddModuleScore(P_01_050, features = ctp[i], name = cnp[i])
P_01_051 = AddModuleScore(P_01_051, features = ctp[i], name = cnp[i])
P_01_051_J10 = AddModuleScore(P_01_051_J10, features = ctp[i], name = cnp[i])
P_01_052 = AddModuleScore(P_01_052, features = ctp[i], name = cnp[i])
P_01_053 = AddModuleScore(P_01_053, features = ctp[i], name = cnp[i])
P_01_054_J0 = AddModuleScore(P_01_054_J0, features = ctp[i], name = cnp[i])
}
##

####

ct= c( NK, B, plasmaCells,Tcells, Platelet )
cn= c(  "NK", "B", "plasmaCells","Tcells", "Platelet")
for (i in seq(1:length(ct))) {
P_01_047= AddModuleScore(P_01_047, features = ct[i], name = cn[i])
P_01_048_J0= AddModuleScore(P_01_048_J0, features = ct[i], name = cn[i])
P_01_048_J5= AddModuleScore(P_01_048_J5, features = ct[i], name = cn[i])
P_01_049 = AddModuleScore(P_01_049, features = ct[i], name = cn[i])
P_01_050 = AddModuleScore(P_01_050, features = ct[i], name = cn[i])
P_01_051 = AddModuleScore(P_01_051, features = ct[i], name = cn[i])
 P_01_051_J10 = AddModuleScore(P_01_051_J10, features = ct[i], name = cn[i])
P_01_052 = AddModuleScore(P_01_052, features = ct[i], name = cn[i])
P_01_053 = AddModuleScore(P_01_053, features = ct[i], name = cn[i])
P_01_054_J0 = AddModuleScore(P_01_054_J0, features = ct[i], name = cn[i])
}


#################################
keepAPC=function(x){
  x=subset(x, pDC1>0.2 | monocytes1>0.2 | cDC1>0.2 | asDC1>0.2) %>% subset(., NK1 <0 & B1 <0 & plasmaCells1 < 0 & Tcells1 < 0 )
  return(x)}

P_01_047= keepAPC(P_01_047) %>%  DietSeurat() %>% NormalizeData()
P_01_048_J0= keepAPC(P_01_048_J0) %>%  DietSeurat(.) %>% NormalizeData()
P_01_048_J5= keepAPC(P_01_048_J5) %>%  DietSeurat() %>% NormalizeData()
P_01_049= keepAPC(P_01_049) %>%  DietSeurat() %>% NormalizeData()
P_01_050= keepAPC(P_01_050) %>%  DietSeurat() %>% NormalizeData()
P_01_051= keepAPC(P_01_051) %>%  DietSeurat() %>% NormalizeData()
P_01_051_J10= keepAPC(P_01_051_J10) %>%  DietSeurat() %>% NormalizeData()
P_01_052= keepAPC(P_01_052) %>%  DietSeurat() %>% NormalizeData()
P_01_053= keepAPC(P_01_053) %>%  DietSeurat() %>% NormalizeData()
P_01_054_J0= keepAPC(P_01_054_J0) %>%  DietSeurat() %>% NormalizeData()

```

```{r}
saveRDS(P_01_047, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/P_01_047_APC.Rds")
saveRDS(P_01_048_J0, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/P_01_048_J0_APC.Rds")
saveRDS(P_01_048_J5, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/P_01_048_J5_APC.Rds")
saveRDS(P_01_049, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/P_01_049_APC.Rds")

saveRDS(P_01_050, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/P_01_050_APC.Rds")
saveRDS(P_01_051_J10, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/P_01_051_J10_APC.Rds")

saveRDS(P_01_052, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/P_01_052_APC.Rds")
saveRDS(P_01_053, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/P_01_053_APC.Rds")
saveRDS(P_01_054_J0, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/P_01_054_J0_APC.Rds")
```


```{r}
tmp1=list.files("C:/Users/MakotoTeam/Documents/scRNAseq_previousData/")

Pr1_05_01_J4=CreateSeuratObject(Read10X("C:/Users/MakotoTeam/Documents/scRNAseq_previousData/05_01_J4/"),min.cells = 20, min.features = 50, project="05_01_J4")

for(i in seq(1, length(tmp1))){
  assign(paste0("Pr1_", tmp1[i]),
         CreateSeuratObject( Read10X(paste0("C:/Users/MakotoTeam/Documents/scRNAseq_previousData/",tmp1[i], "/")), project = tmp1[i],min.cells = 20, min.features = 50))}

```

Add the mito genes ratio:
```{r}

nm=grep("Pr1_",ls(),value = T)
covid_round1=list(Pr1_01_040_J2,Pr1_01_042_J4)
####

   Pr1_01_040_J2[["percent.mito"]] <- PercentageFeatureSet( Pr1_01_040_J2, pattern = "^MT-")
    Pr1_01_042_J4[["percent.mito"]] <- PercentageFeatureSet( Pr1_01_042_J4, pattern = "^MT-")
     Pr1_01_37_J26[["percent.mito"]] <- PercentageFeatureSet( Pr1_01_37_J26, pattern = "^MT-")
      Pr1_01_42_J0[["percent.mito"]] <- PercentageFeatureSet( Pr1_01_42_J0, pattern = "^MT-")
      ##
       Pr1_01_42_J16[["percent.mito"]] <- PercentageFeatureSet( Pr1_01_42_J16, pattern = "^MT-")
        Pr1_01_43_J0[["percent.mito"]] <- PercentageFeatureSet( Pr1_01_43_J0, pattern = "^MT-")
        
         Pr1_03_001_J1[["percent.mito"]] <- PercentageFeatureSet( Pr1_03_001_J1, pattern = "^MT-")
         Pr1_03_01_J21[["percent.mito"]] <- PercentageFeatureSet( Pr1_03_01_J21, pattern = "^MT-")
           Pr1_03_01_J5[["percent.mito"]] <- PercentageFeatureSet( Pr1_03_01_J5, pattern = "^MT-")
            Pr1_04_001_J5[["percent.mito"]] <- PercentageFeatureSet( Pr1_04_001_J5, pattern = "^MT-")
            ##
             Pr1_04_01_J0[["percent.mito"]] <- PercentageFeatureSet( Pr1_04_01_J0, pattern = "^MT-")
              Pr1_05_01_J1[["percent.mito"]] <- PercentageFeatureSet( Pr1_05_01_J1, pattern = "^MT-")
               Pr1_05_01_J4[["percent.mito"]] <- PercentageFeatureSet( Pr1_05_01_J4, pattern = "^MT-")
               ##
               Pr1_05_02[["percent.mito"]] <- PercentageFeatureSet( Pr1_05_02, pattern = "^MT-")
             
                Pr1_CTRL_06_01_EDTA[["percent.mito"]] <- PercentageFeatureSet( Pr1_CTRL_06_01_EDTA, pattern = "^MT-")
                 Pr1_CTRL_RPMI[["percent.mito"]] <- PercentageFeatureSet( Pr1_CTRL_RPMI, pattern = "^MT-")
                  Pr1_P037_J0[["percent.mito"]] <- PercentageFeatureSet( Pr1_P037_J0, pattern = "^MT-")
                   Pr1_P037_J4[["percent.mito"]] <- PercentageFeatureSet( Pr1_P037_J4, pattern = "^MT-")
                   
#######
    Pr1_01_040_J2 <- QC( Pr1_01_040_J2)
    Pr1_01_042_J4 <- QC( Pr1_01_042_J4)
     Pr1_01_37_J26 <- QC( Pr1_01_37_J26)
      Pr1_01_42_J0 <- QC( Pr1_01_42_J0)
      ##
       Pr1_01_42_J16 <- QC( Pr1_01_42_J16)
        Pr1_01_43_J0 <- QC( Pr1_01_43_J0)
        
         Pr1_03_001_J1 <- QC( Pr1_03_001_J1)
         Pr1_03_01_J21 <- QC( Pr1_03_01_J21)
           Pr1_03_01_J5 <- QC( Pr1_03_01_J5)
            Pr1_04_001_J5 <- QC( Pr1_04_001_J5)
            ##
             Pr1_04_01_J0 <- QC( Pr1_04_01_J0)
              Pr1_05_01_J1 <-QC( Pr1_05_01_J1)
              Pr1_05_01_J4 <-QC( Pr1_05_01_J4)
               Pr1_05_02<- QC( Pr1_05_02)
                Pr1_CTRL_06_01_EDTA <- QC( Pr1_CTRL_06_01_EDTA)
                 Pr1_CTRL_RPMI <- QC( Pr1_CTRL_RPMI)
                  Pr1_P037_J0 <- QC( Pr1_P037_J0)
                   Pr1_P037_J4<- QC( Pr1_P037_J4)
  ######################################
       Pr1_01_040_J2 <- NormalizeData( Pr1_01_040_J2) %>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
    Pr1_01_042_J4 <- NormalizeData( Pr1_01_042_J4)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
     Pr1_01_37_J26 <- NormalizeData( Pr1_01_37_J26)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
      Pr1_01_42_J0 <- NormalizeData( Pr1_01_42_J0)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
      ##
       Pr1_01_42_J16 <- NormalizeData( Pr1_01_42_J16)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
        Pr1_01_43_J0 <- NormalizeData( Pr1_01_43_J0)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
        
         Pr1_03_001_J1 <- NormalizeData( Pr1_03_001_J1)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
         Pr1_03_01_J21 <- NormalizeData( Pr1_03_01_J21)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
           Pr1_03_01_J5 <- NormalizeData( Pr1_03_01_J5)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
            Pr1_04_001_J5 <- NormalizeData( Pr1_04_001_J5)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
            ##
             Pr1_04_01_J0 <- NormalizeData( Pr1_04_01_J0)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
              Pr1_05_01_J1 <-NormalizeData( Pr1_05_01_J1)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
               Pr1_05_01_J4 <-NormalizeData( Pr1_05_01_J4)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
               
               Pr1_05_02<- NormalizeData( Pr1_05_02)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
                Pr1_CTRL_06_01_EDTA <- NormalizeData( Pr1_CTRL_06_01_EDTA)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
                 Pr1_CTRL_RPMI <- NormalizeData( Pr1_CTRL_RPMI)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
                  Pr1_P037_J0 <- NormalizeData( Pr1_P037_J0)%>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)
                   Pr1_P037_J4<- NormalizeData( Pr1_P037_J4)   %>% FindVariableFeatures(., selection.method = "vst", nfeatures = 3000)          


 ########################
                   
ctp= c( pDC, cDC, monocytes, asDC)
cnp= c( "pDC", "cDC", "monocytes", "asDC")

for (i in seq(1:length(ctp))) {
  Pr1_05_01_J4 =AddModuleScore( Pr1_05_01_J4, features = ctp[i], name = cnp[i])
  
   Pr1_01_040_J2 <- AddModuleScore( Pr1_01_040_J2, features = ctp[i], name = cnp[i])
    Pr1_01_042_J4 <- AddModuleScore( Pr1_01_042_J4, features = ctp[i], name = cnp[i])
     Pr1_01_37_J26 <- AddModuleScore( Pr1_01_37_J26, features = ctp[i], name = cnp[i])
      Pr1_01_42_J0 <- AddModuleScore( Pr1_01_42_J0, features = ctp[i], name = cnp[i])
      ##
       Pr1_01_42_J16 <- AddModuleScore( Pr1_01_42_J16, features = ctp[i], name = cnp[i])
        Pr1_01_43_J0 <- AddModuleScore( Pr1_01_43_J0, features = ctp[i], name = cnp[i])
        
         Pr1_03_001_J1 <- AddModuleScore( Pr1_03_001_J1, features = ctp[i], name = cnp[i])
         Pr1_03_01_J21 <- AddModuleScore( Pr1_03_01_J21, features = ctp[i], name = cnp[i])
           Pr1_03_01_J5 <- AddModuleScore( Pr1_03_01_J5, features = ctp[i], name = cnp[i])
            Pr1_04_001_J5 <- AddModuleScore( Pr1_04_001_J5, features = ctp[i], name = cnp[i])
            ##
             Pr1_04_01_J0 <- AddModuleScore( Pr1_04_01_J0, features = ctp[i], name = cnp[i])
              Pr1_05_01_J1 <-AddModuleScore( Pr1_05_01_J1, features = ctp[i], name = cnp[i])
               Pr1_05_02<- AddModuleScore( Pr1_05_02, features = ctp[i], name = cnp[i])
                Pr1_CTRL_06_01_EDTA <- AddModuleScore( Pr1_CTRL_06_01_EDTA, features = ctp[i], name = cnp[i])
                 Pr1_CTRL_RPMI <- AddModuleScore( Pr1_CTRL_RPMI, features = ctp[i], name = cnp[i])
                  Pr1_P037_J0 <- AddModuleScore( Pr1_P037_J0, features = ctp[i], name = cnp[i])
                   Pr1_P037_J4<- AddModuleScore( Pr1_P037_J4, features = ctp[i], name = cnp[i])}


####

ct= c( NK, B, plasmaCells,Tcells, Platelet )
cn= c(  "NK", "B", "plasmaCells","Tcells", "Platelet")
for (i in seq(1:length(ct))) {
  Pr1_05_01_J4 =AddModuleScore( Pr1_05_01_J4, features = ctp[i], name = cnp[i])

   Pr1_05_01_J4 =AddModuleScore( Pr1_05_01_J4, features = ctp[i], name = cnp[i])
 Pr1_01_040_J2 <- AddModuleScore( Pr1_01_040_J2, features = ct[i], name = cn[i])
    Pr1_01_042_J4 <- AddModuleScore( Pr1_01_042_J4, features = ct[i], name = cn[i])
     Pr1_01_37_J26 <- AddModuleScore( Pr1_01_37_J26, features = ct[i], name = cn[i])
      Pr1_01_42_J0 <- AddModuleScore( Pr1_01_42_J0, features = ct[i], name = cn[i])
      ##
       Pr1_01_42_J16 <- AddModuleScore( Pr1_01_42_J16, features = ct[i], name = cn[i])
        Pr1_01_43_J0 <- AddModuleScore( Pr1_01_43_J0, features = ct[i], name = cn[i])
        
         Pr1_03_001_J1 <- AddModuleScore( Pr1_03_001_J1, features = ct[i], name = cn[i])
         Pr1_03_01_J21 <- AddModuleScore( Pr1_03_01_J21, features = ct[i], name = cn[i])
           Pr1_03_01_J5 <- AddModuleScore( Pr1_03_01_J5, features = ct[i], name = cn[i])
            Pr1_04_001_J5 <- AddModuleScore( Pr1_04_001_J5, features = ct[i], name = cn[i])
            ##
             Pr1_04_01_J0 <- AddModuleScore( Pr1_04_01_J0, features = ct[i], name = cn[i])
              Pr1_05_01_J1 <-AddModuleScore( Pr1_05_01_J1, features = ct[i], name = cn[i])
               Pr1_05_02<- AddModuleScore( Pr1_05_02, features = ct[i], name = cn[i])
                Pr1_CTRL_06_01_EDTA <- AddModuleScore( Pr1_CTRL_06_01_EDTA, features = ct[i], name = cn[i])
                 Pr1_CTRL_RPMI <- AddModuleScore( Pr1_CTRL_RPMI, features = ct[i], name = cn[i])
                  Pr1_P037_J0 <- AddModuleScore( Pr1_P037_J0, features = ct[i], name = cn[i])
                   Pr1_P037_J4<- AddModuleScore( Pr1_P037_J4, features = ct[i], name = cn[i])}
########

 Pr1_01_040_J2 <- AddModuleScore( Pr1_01_040_J2, features = asDC, name = "asDC")
    Pr1_01_042_J4 <- AddModuleScore( Pr1_01_042_J4, features = asDC, name = "asDC")
     Pr1_01_37_J26 <- AddModuleScore( Pr1_01_37_J26, features = asDC, name = "asDC")
      Pr1_01_42_J0 <- AddModuleScore( Pr1_01_42_J0, features = asDC, name = "asDC")
      ##
       Pr1_01_42_J16 <- AddModuleScore( Pr1_01_42_J16, features = asDC, name = "asDC")
        Pr1_01_43_J0 <- AddModuleScore( Pr1_01_43_J0, features = asDC, name = "asDC")
        
         Pr1_03_001_J1 <- AddModuleScore( Pr1_03_001_J1, features = asDC, name = "asDC")
         Pr1_03_01_J21 <- AddModuleScore( Pr1_03_01_J21, features = asDC, name = "asDC")
           Pr1_03_01_J5 <- AddModuleScore( Pr1_03_01_J5, features = asDC, name = "asDC")
            Pr1_04_001_J5 <- AddModuleScore( Pr1_04_001_J5, features = asDC, name = "asDC")
            ##
             Pr1_04_01_J0 <- AddModuleScore( Pr1_04_01_J0, features = asDC, name = "asDC")
              Pr1_05_01_J1 <-AddModuleScore( Pr1_05_01_J1, features = asDC, name = "asDC")
               Pr1_05_02<- AddModuleScore( Pr1_05_02, features = asDC, name = "asDC")
                Pr1_CTRL_06_01_EDTA <- AddModuleScore( Pr1_CTRL_06_01_EDTA, features = asDC, name = "asDC")
                 Pr1_CTRL_RPMI <- AddModuleScore( Pr1_CTRL_RPMI, features = asDC, name = "asDC")
                  Pr1_P037_J0 <- AddModuleScore( Pr1_P037_J0, features = asDC, name = "asDC")
                   Pr1_P037_J4<- AddModuleScore( Pr1_P037_J4, features = asDC, name = "asDC")
                   Pr1_01_040_J2$asDC1=rep(0, ncol(Pr1_01_040_J2))
###############

 Pr1_01_040_J2 <- keepAPC( Pr1_01_040_J2)%>%  DietSeurat() %>% NormalizeData()

    Pr1_01_042_J4 <- keepAPC( Pr1_01_042_J4)%>%  DietSeurat() %>% NormalizeData()
     Pr1_01_37_J26 <- keepAPC( Pr1_01_37_J26)%>%  DietSeurat() %>% NormalizeData()
      Pr1_01_42_J0 <- keepAPC( Pr1_01_42_J0)%>%  DietSeurat() %>% NormalizeData()
      ##
       Pr1_01_42_J16 <- keepAPC( Pr1_01_42_J16)%>%  DietSeurat() %>% NormalizeData()
        Pr1_01_43_J0 <- keepAPC( Pr1_01_43_J0)%>%  DietSeurat() %>% NormalizeData()
        
         Pr1_03_001_J1 <- keepAPC( Pr1_03_001_J1)%>%  DietSeurat() %>% NormalizeData()
         Pr1_03_01_J21 <- keepAPC( Pr1_03_01_J21)%>%  DietSeurat() %>% NormalizeData()
           Pr1_03_01_J5 <- keepAPC( Pr1_03_01_J5)%>%  DietSeurat() %>% NormalizeData()
            Pr1_04_001_J5 <- keepAPC( Pr1_04_001_J5)%>%  DietSeurat() %>% NormalizeData()
            ##
             Pr1_04_01_J0 <- keepAPC( Pr1_04_01_J0)%>%  DietSeurat() %>% NormalizeData()
              Pr1_05_01_J1 <-keepAPC( Pr1_05_01_J1)%>%  DietSeurat() %>% NormalizeData()
               Pr1_05_01_J4 <-keepAPC( Pr1_05_01_J4)%>%  DietSeurat() %>% NormalizeData()
               Pr1_05_02<- keepAPC( Pr1_05_02)%>%  DietSeurat() %>% NormalizeData()
                Pr1_CTRL_06_01_EDTA <- keepAPC( Pr1_CTRL_06_01_EDTA)%>%  DietSeurat() %>% NormalizeData()
                 Pr1_CTRL_RPMI <- keepAPC( Pr1_CTRL_RPMI)%>%  DietSeurat() %>% NormalizeData()
                  Pr1_P037_J0 <- keepAPC( Pr1_P037_J0)%>%  DietSeurat() %>% NormalizeData()
                   Pr1_P037_J4<- keepAPC( Pr1_P037_J4)%>%  DietSeurat() %>% NormalizeData()
```

save the objects:
```{r}

saveRDS(Pr1_01_040_J2, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/Pr1_01_040_J2.Rds"  )
saveRDS( Pr1_01_042_J4 ,"C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/Pr1_01_042_J4.Rds")
saveRDS( Pr1_01_37_J26 , "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_01_37_J26.Rds")
saveRDS( Pr1_01_42_J0, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/Pr1_01_42_J0.Rds")
      ##
saveRDS( Pr1_01_42_J16 ,"C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/Pr1_01_42_J16.Rds")
saveRDS(  Pr1_01_43_J0, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_01_43_J0.Rds")
        
saveRDS(  Pr1_03_001_J1,"C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_03_001_J1.Rds")
saveRDS(  Pr1_03_01_J21, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_03_01_J21.Rds")
saveRDS(  Pr1_03_01_J5, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_03_01_J5.Rds")
saveRDS( Pr1_04_001_J5, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_04_001_J5.Rds")
            ##
saveRDS( Pr1_04_01_J0, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_04_01_J0.Rds")
saveRDS(  Pr1_05_01_J1, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_05_01_J1.Rds")
saveRDS(  Pr1_05_01_J4, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_05_01_J4.Rds")
##
saveRDS(   Pr1_05_02,"C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_05_02.Rds")
saveRDS( Pr1_CTRL_06_01_EDTA, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_CTRL_06_01_EDTA.Rds")
saveRDS( Pr1_CTRL_RPMI ,"C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_CTRL_RPMI.Rds")
saveRDS(  Pr1_P037_J0 ,"C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_P037_J0.Rds")
saveRDS(  Pr1_P037_J4,"C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/APC_CTRL_COVID_ROUND1/ Pr1_P037_J4.Rds")
```


```{r}
library(devtools)
install_github("immunogenomics/harmony")
library(harmony)
```


RUN INTEGRATION:
```{r}
COVID.list <-list(P_01_047,P_01_048_J0,P_01_048_J5,P_01_049,P_01_050,P_01_051,P_01_051_J10,P_01_052,P_01_053, P_01_054_J0, c1, c2,c3, Pr1_01_040_J2,Pr1_01_042_J4,Pr1_01_37_J26,Pr1_01_42_J0,Pr1_01_42_J16,Pr1_01_43_J0, Pr1_03_001_J1,Pr1_03_01_J21,Pr1_03_01_J5, Pr1_04_001_J5,Pr1_04_01_J0,Pr1_05_01_J1,Pr1_05_01_J4,Pr1_05_02,Pr1_CTRL_06_01_EDTA,Pr1_CTRL_RPMI,Pr1_P037_J0,Pr1_P037_J4 )
####
QC2=function(x){
  x=subset(x, subset  = nFeature_RNA > 100 & nCount_RNA > 100 )
  return(x)}

for (i in 1:length(COVID.list)) {
   # COVID.list[[i]] <- QC2(COVID.list[[i]])
      COVID.list[[i]]=subset( COVID.list[[i]], subset  = nFeature_RNA > 100 & nCount_RNA > 100 )}


library("glmGamPoi")

k.filter <- min(200, min(sapply(COVID.list, ncol)))

for (i in 1:length(COVID.list)) {
    COVID.list[[i]] <- SCTransform(COVID.list[[i]], verbose = FALSE,variable.features.n = 2000,method = "glmGamPoi")}

COVID.features <- SelectIntegrationFeatures(object.list = COVID.list, nfeatures = 3000,k.filter = k.filter)

COVID_merged <- merge(COVID.list[[1]], y = COVID.list[2:length(COVID.list)],   project = "COVID",  merge.data = TRUE)
VariableFeatures(COVID_merged) <- COVID.features

COVID_merged <- RunPCA(object = COVID_merged, assay = "SCT", npcs = 50)
COVID_merged <- RunHarmony(object = COVID_merged,
                                    assay.use = "SCT",
                                    reduction = "pca",
                                    group.by.vars = "orig.ident",
                                    dims.use = 1:50,
                                    plot_convergence = TRUE)
COVID_merged <- RunUMAP(object = COVID_merged, assay = "SCT", reduction = "harmony", dims = 1:50)
COVID_merged <- FindNeighbors(object = COVID_merged, assay = "SCT", reduction = "harmony", dims = 1:50)
DefaultAssay(COVID_merged) = "SCT"
COVID_merged <- FindClusters(object = COVID_merged, resolution = 1.2)
```


ADD ANNOTATION:
```{r}
DefaultAssay(COVID_merged)="RNA"
VlnPlot(COVID_merged, c("TCF4", "CLEC4C", "IRF7", "IRF8", "LILRA4", "IL3RA", "TLR9", "SPIB"))
VlnPlot(COVID_merged, c("ANPEP",  "CD1C", "ITGAX", "CST3", "FCER1A"))
VlnPlot(COVID_merged, c("CD14", "FCGR1A"," S100A12", "FCGR3A", "MS4A7", "LYZ", "CXCR3"))
VlnPlot(COVID_merged, c("AXL", "SIGLEC6","CD22"))
```

CORRECTED ON 08/12:
Increased clustering resolution to 1.0

```{r}
Idents(COVID_merged)="seurat_clusters"
cellannotation= c("CD14+ Monocytes","CD1c+ DC",
                  "CD14+ Monocytes","CD14+ Monocytes","CD14+ Monocytes",
                  "pDC", "CD14+ Monocytes", "CD14+ Monocytes",
                  "CD1c+ DC", "CD14+ Monocytes",
                  "CD16+ Monocytes","CLEC9a+ DC",
                  "CD14+ Monocytes", "CD14+ Monocytes","CD14+ Monocytes",
                  "asDC",
                  "CD14+ Monocytes","Undetermined",
                  "CD1c+ DC","CD1c+ DC",  "CD1c+ DC", "CD14+ Monocytes","CD14+ Monocytes",
                   "CD1c+ DC", "Undetermined", "Undetermined", "Undetermined")
Idents(COVID_merged)="seurat_clusters"
                  
names(cellannotation)= levels(Idents(COVID_merged))

COVID_merged=RenameIdents(COVID_merged, cellannotation)
COVID_merged$celltype=Idents(COVID_merged)
covid= COVID_merged
###remove the undetermined:
COVID_merged= SubsetData(COVID_merged, ident.remove = "Undetermined")
saveRDS(COVID_merged, "C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/COVID_EDTA_RPMI_RDS/COVID_Merged_12122020.Rds")
```

```{r}
path="C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/COVID_EDTA_RPMI_RDS/"
library(RColorBrewer)
library(ggpubr)
getPalette = colorRampPalette(brewer.pal(6, "Set1"))
p1=DimPlot(COVID_merged, group.by = "celltype", cols= getPalette(length(unique(COVID_merged$celltype))), label.size = 6, pt.size = 0.8, label.color = "black", label.box = T) 
##
ggsave(filename=paste0("improved_COVID19_ALL_merged_APC_UMAP_CELLTYPE_",Sys.Date(),".png"), plot=p1, device="png",
       path= path, height=8, width=8, units="in", dpi=500)
##
getPalette = colorRampPalette(brewer.pal(8, "Dark2"))
p2=DimPlot(COVID_merged, group.by = "orig.ident", cols= getPalette(length(unique(COVID_merged$orig.ident))),label.size = 6, pt.size = 0.8)

ggsave(filename=paste0("improved_COVID19_ALL_merged_APC_UMAP_Patient_",Sys.Date(),".png"), plot=p2, device="png",
       path= path, height=8, width=8, units="in", dpi=500)
```


ADD THE METADATA ON SEVERITY AND ON EDTA VS RPMI:
ADD EXPR CONDITION:
```{r}
nm=grep("Pr1_",ls(),value = T)

COVID_merged$condition =
  ifelse(COVID_merged$orig.ident== "05_01_J1"  | COVID_merged$orig.ident== "05_01_J4"| COVID_merged$orig.ident== "05_02" | COVID_merged$orig.ident== "CTRL_06_01_EDTA" | COVID_merged$orig.ident== "P037_J0" | COVID_merged$orig.ident== "P037_J4" | COVID_merged$orig.ident== "03_001_J1" | COVID_merged$orig.ident== "03_01_J5"| COVID_merged$orig.ident== "01_040_J2"| 
           
           COVID_merged$orig.ident== "CTRL_02021" |  COVID_merged$orig.ident== "CTRL_02022" |  COVID_merged$orig.ident== "01_049" |  COVID_merged$orig.ident== "01_052" |  COVID_merged$orig.ident== "01_054_J0" , "EDTA", "RPMI")
```

when annotating do not forget to respect the same nomenclature
```{r}
COVID_merged$severity =
  ifelse(COVID_merged$orig.ident== "01_052"  | COVID_merged$orig.ident== "01_051"| COVID_merged$orig.ident== "01_048_J0" | COVID_merged$orig.ident== "01_048_J5" | COVID_merged$orig.ident== "04_001_J5" | COVID_merged$orig.ident== "04_01_J0" | COVID_merged$orig.ident== "05_01_J4" | COVID_merged$orig.ident== "05_01_J1" | COVID_merged$orig.ident== "05_02" , "Moderate",
   ifelse( COVID_merged$orig.ident== "CTRL_RPMI"| COVID_merged$orig.ident== "CTRL_02020RPMI"| COVID_merged$orig.ident== "CTRL_02021"| COVID_merged$orig.ident== "CTRL_02022"| COVID_merged$orig.ident== "CTRL_06_01_EDTA", "Healthy", "Severe"))
```


```{r}
female= WhichCells(COVID_merged, expression =  XIST > 0.1)
COVID_merged$gender= ifelse(colnames(COVID_merged) %in% female, "Female", "Male" ) 
p= DimPlot(COVID_merged, group.by  = "gender")

ggsave(filename=paste0("improved_COVID19_ALL_merged_APC_UMAP_Gender_",Sys.Date(),".png"), plot=p, device="png",
       path= path, height=8, width=8, units="in", dpi=500)
```



```{r}
getPalette = colorRampPalette(brewer.pal(6, "Dark2"))
p3=DimPlot(COVID_merged, group.by = "severity", cols= getPalette(length(unique(COVID_merged$severity))), label.size = 6, pt.size = 0.5, label.color = "black", label.box = T) 

ggsave(filename=paste0("improved_COVID19_ALL_merged_APC_UMAP_Severity_2_",Sys.Date(),".png"), plot=p3, device="png",
       path= path, height=8, width=8, units="in", dpi=500)
####
getPalette = colorRampPalette(brewer.pal(7, "Dark2"))

p4=DimPlot(COVID_merged, group.by = "condition", cols= getPalette(length(unique(COVID_merged$condition))), label.size = 6, pt.size = 0.5, label.color = "black", label.box = T) 

ggsave(filename=paste0("improvedCOVID19_ALL_merged_APC_UMAP_CONDITION_2_",Sys.Date(),".png"), plot=p4, device="png",
       path= path, height=8, width=8, units="in", dpi=500)
```

Generate the numbers: per edta and rpmi

```{r}
countable=function(c1, c2){
xd=table(c1, c2)
a=colnames(xd)
b=rownames(xd)
xd=as.matrix.data.frame(xd)
colnames(xd)=a
rownames(xd)=b
xd=as.data.frame(xd)
return(xd)}

s1=countable(COVID_merged$orig.ident, COVID_merged$celltype)
s2= countable( COVID_merged$celltype,COVID_merged$condition)
##

write.csv(s1, paste0(path,"improved_CELLTYPE_PATIENT.csv"))
write.csv(s2, paste0(path,"improved_CELLTYPE_CONDITION.csv"))
```

```{r}
saveRDS(COVID_merged,paste0("C:/Users/MakotoTeam/Documents/scRNAseq_Round3/Analysis_28Novembre/COVID_EDTA_RPMI_RDS/","COVID_MERGED_EDTA_RPMI_CORRECTED_", Sys.Date(), ".Rds"))
```




